Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}

### 2019-09-04 paper/tex/analysis.tex
incorporated feedback from @morriscb

The exact implementation of the metrics theoretically depends on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:2018}.
Even considering a single method under the same parametrization, such as the 200-bin $0 < z < 2$ piecewise constant function used here, the exact bin definitions will affect the result.
%\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of \pzpdf s for the final \lsst\ Project tables.\footnote{See, e.~g.~the \lsst\ Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
We will discuss the choice of \pzpdf\ parameterization further in Section~\ref{sec:discussion}.

@@

Because \lsst's \pzpdf s will be used for many scientific applications, some of which require accuracy of each individual catalog entry, we consider several metrics that probe the population-level performance of the \pzpdf s.
Because we have the true redshifts but not true \pzpdf s for comparison, we remind the reader of the Cumulative Distribution Function (CDF)
\begin{equation}
  \label{eq:cdf}
  \mathrm{CDF}[f, q] \equiv \int_{-\infty}^{q} f(z) dz,
\end{equation}
of a generic univariate PDF $f(z)$, which is used as the basis for several of our metrics.

@@

@@

The \buzz\ simulation process precludes testing the degree to which samples from our \pz\ posteriors reconstruct the space of $p(z, \mathrm{data})$.
To the knowledge of the authors, there is only one metric that can be used to evaluate the performance of individual \pzpdf\ estimators in the absence of true \pz\ posteriors.
Using the notation introduced in Section~\ref{sec:flexzboost}, the conditional density estimation (CDE) loss is defined as
\begin{equation}
  \label{eq:cde-loss}
  L(f, \widehat{f}) \equiv \int \int (f(z \vert \x) - \widehat{f}(z \vert \x))^{2} \mathrm{d}z \mathrm{d}P(\x)
\end{equation}
in terms of the photometry $\x$, an analogue to the familiar root-mean-square-error used in conventional regression.
We estimate the CDE loss via
\begin{equation}
  \label{eq:estimated-cde-loss}
  \hat{L}(f, \widehat{f}) = \mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X, Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
\end{equation}
where the first term is the expectation value of the \pz\ posterior with respect to the marginal distribution of the photometric covariates $\X$, the second term is the expectation value with respect to the joint distribution of $\X$ and the space $Z$ of all possible redshifts, and the third term $K_{f}$ is a constant depending only upon the true conditional densities $f(z \vert \x)$.
We may estimate these expectations empirically on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.

>>>>

The exact implementation of the metrics theoretically depends on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:2018}.
Even considering a single method under the same parametrization, such as the 200-bin $0 < z < 2$ piecewise constant function used here, the exact bin definitions must affect the result.
%\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of \pzpdf s for the final \lsst\ Project tables.\footnote{See, e.~g.~the \lsst\ Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
We will discuss the choice of \pzpdf\ parameterization further in Section~\ref{sec:discussion}.

@@

Because \lsst's \pzpdf s will be used for many scientific applications, some of which require accuracy of each individual catalog entry, we consider several metrics that probe the population-level performance of the \pzpdf s.
As we have the true redshifts but not true \pzpdf s for comparison, we remind the reader of the Cumulative Distribution Function (CDF)
\begin{equation}
  \label{eq:cdf}
  \mathrm{CDF}[f, q] \equiv \int_{-\infty}^{q} f(z) dz,
\end{equation}
of a generic univariate PDF $f(z)$, which is used as the basis for several of our metrics.
We describe metrics based on the CDF in Section~\ref{sec:qqpit} and metrics of summary statistics thereof in Section~\ref{sec:summqqpit}.

\subsubsection{CDF-based metrics}
\label{sec:qqpit}

@@

\subsubsection{Summary statistics of CDF-based metrics}
\label{sec:summqqpit}

@@

The \buzz\ simulation process precludes testing the degree to which samples from our \pz\ posteriors reconstruct the space of $p(z, \mathrm{data})$.
To the knowledge of the authors, there is only one metric that can be used to evaluate the performance of individual \pzpdf\ estimators in the absence of true \pz\ posteriors.
The conditional density estimation (CDE) loss is an analogue to the familiar root-mean-square-error used in conventional regression, defined as
\begin{equation}
  \label{eq:cde-loss}
  L(f, \widehat{f}) \equiv \int \int (f(z \vert \x) - \widehat{f}(z \vert \x))^{2} \mathrm{d}z \mathrm{d}P(\x) ,
\end{equation}
where $f(z \vert \x)$ is the true \pzpdf\ that we do not have and $\widehat{f}(z \vert \x)$ is an estimate thereof, in terms of the photometry $\x$.
(See Section~\ref{sec:flexzboost} for a review of the notation.)
We estimate the CDE loss via
\begin{equation}
  \label{eq:estimated-cde-loss}
  \hat{L}(f, \widehat{f}) = \mathbb{E}_\X \left[\int \widehat{f}(z \mid \X)^{2} dz\right] - 2 \mathbb{E}_{\X, Z}\left[\widehat{f}(Z \mid \X)\right] + K_{f},
\end{equation}
where the first term is the expectation value of the \pz\ posterior with respect to the marginal distribution of the photometric covariates $\X$, the second term is the expectation value with respect to the joint distribution of $\X$ and the space $Z$ of all possible redshifts, and the third term $K_{f}$ is a constant depending only upon the true conditional densities $f(z \vert \x)$.
We may estimate these expectations empirically on the test or validation data \citep[Eq.~7 in][]{Izbicki:17b} without knowledge of the true densities.

### 2019-09-04 paper/tex/discussion.tex
incorporated feedback from @morriscb

The \trainz\ estimator of Section~\ref{sec:trainz}, which assigns every galaxy a \pzpdf\ equal to $N(z)$ of the training set, is introduced as an experimental control or null test to demonstrate this point via \textit{reductio ad absurdum}.
Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
\textbf{We make the alarming observation that \trainz, the experimental control, outperforms all codes on the CDF-based metrics, and all but one code on the $N(z)$ based statistics.}
The PIT and other CDF-based metrics upon which modern \pzpdf\ comparisons are built can be gamed by a trivial estimator that yields only an affirmation of prior knowledge uninformed by the data.
In other words, such ensemble metrics are not appropriate for the task of selecting \pzpdf\ codes for analysis pipelines.
%\textit{Explain the connection between \trainz\ and these metrics.}

>>>>

The \trainz\ estimator of Section~\ref{sec:trainz}, which assigns every galaxy a \pzpdf\ equal to $N(z)$ of the training set, is introduced as an experimental control or null test to demonstrate this point via \textit{reductio ad absurdum}.
Because our training set is perfectly representative of the test set, $N(z)$ should be identical for both sets down to statistical noise.
\textbf{We make the alarming observation that \trainz, the experimental control, outperforms all codes on the CDF-based metrics, and all but one code on the $N(z)$ based statistics.}
The PIT and other CDF-based metrics upon which modern \pzpdf\ comparisons are built \citep{Bordoloi:10,Polsterer:16,Tanaka:17} can be gamed by a trivial estimator that yields only an affirmation of prior knowledge uninformed by the data.
In other words, such ensemble metrics are not appropriate for the task of selecting \pzpdf\ codes for analysis pipelines.
%\textit{Explain the connection between \trainz\ and these metrics.}

### 2019-09-04 paper/tex/introduction.tex
incorporated feedback from @morriscb

The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\proj{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\proj{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\proj{HSC},][]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\proj{WFIRST},][]{Green:12}, represent a paradigm shift to reliance on photometric, rather than solely spectroscopic, galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts ($z$).

Effective astrophysical inference using the catalogues resulting from these ongoing and upcoming missions, however, necessitates accurate and precise photometric redshift (\pz) estimation methodologies.
As an example, in order for \pz\ systematics to not dominate the statistical noise floor of \lsst's main cosmological sample of $\sim 10^{7}$ galaxies, the \lsst\ Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} specifies that individual galaxy \pz s must have root-mean-square error $\sigma_z < 0.02 (1+z)$, $3 \sigma$ catastrophic outlier rate below $10\%$, and bias below $0.003$.
Specific science cases may have their own requirements on \pz\ performance that exceed those of the survey as a whole.
In that vein, the \lsst\ Dark Energy Science Collaboration (\lsstdesc) developed a separate SRD \citep{Mandelbaum:2018} that conservatively forecasts the constraining power of five cosmological probes, leading to even more stringent requirements on \pz\ performance, including those defined in terms of tomographically binned subsamples populations rather than individual galaxies.

@@

Though the standard has long been for each galaxy in a photometric catalogue to have a \pz\ point estimate and Gaussian error bar, even early applications of \pz s in precision cosmology indicate the inadequacy of point estimates \citep{mandelbaum:2008} to encapsulate the degeneracies resulting from the nontrivial mapping between broad band fluxes and redshift.
Far from a hypothetical situation, this degeneracy is a real consequence of the same deep imaging that enables larger galaxy catalogue sizes.
The lower luminosity and higher redshift populations captured by deeper imaging introduce major physical systematics to \pz s, among them the Lyman break/Balmer break degeneracy, that did not affect shallower large area surveys like the Sloan Digital Sky Survey \citep[\textsc{SDSS},][]{York:00} and Two Micron All Sky Survey \citep[\textsc{2MASS},][]{Skrutskie:06}.

@@

There are numerous techniques for deriving \pzpdf s, yet no one method has been established as clearly superior.
In fact, in the absence of simulated data drawn from known redshift distributions, the very concept of a ``true PDF'' for an individual galaxy is unavailable, and we must instead rely on measures of ensemble behaviour to characterize PDF quality (see \S\,\ref{sec:metrics} for further discussion).
This caveat aside, quantitative comparisons of \pz\ methods have been made before; the \Pz\ Accuracy And Testing \citep[\textsc{PHAT},][]{Hildebrandt:10} effort focused on \pz\ point estimates derived from many photometric bands.
\citet{Rau:2015} introduced a new method for improving \pzpdf s using an ordinal classification algorithm.
\textsc{DES} compared several codes for \pz\ point estimates and a subset with \pzpdf\ information \citep{Sanchez:14} and examined summary statistics of \pzpdf s for tomographically binned galaxy subsamples \citep{Bonnett:16}.

This paper is distinguished from other comparisons of \pz\ methods by its focus on the evaluation criteria for \pzpdf s and interpretation thereof.
We aim to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to ultimately select those that will become part of the \lsst\ pipelines, as part of a key project of the Photometric Redshifts working group of the \lsstdesc, described in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
In this initial study, we focus on evaluating the performance of \pzpdf\ codes using PDF-specific performance metrics in a controlled experiment with complete and representative prior information (template libraries and training sets) to set a baseline for subsequent investigations.
This approach probes how each code considered exploits the information content of the data versus prior information from template libraries and training sets.

>>>>

The current and next generations of large-scale galaxy surveys, including the Dark Energy Survey \citep[\proj{DES},][]{Abbott:05}, the Kilo-Degree Survey \citep[\proj{KiDS},][]{de_Jong:13}, Hyper Suprime-Cam Survey \citep[\proj{HSC},][]{Aihara:2018a,Aihara:2018b}, Large Synoptic Survey Telescope \citep[\lsst,][]{Abell:09}, Euclid \citep{Laureijs:11}, and Wide-Field Infrared Survey Telescope \citep[\proj{WFIRST},][]{Green:12}, represent a paradigm shift to reliance on photometric, rather than solely spectroscopic, galaxy catalogues of substantially larger size at a cost of lacking complete spectroscopically confirmed redshifts ($z$).
Effective astrophysical inference using the catalogues resulting from these ongoing and upcoming missions, however, necessitates accurate and precise photometric redshift (\pz) estimation methodologies.

As an example, in order for \pz\ systematics to not dominate the statistical noise floor of \lsst's main cosmological sample of $\sim 10^{7}$ galaxies, the \lsst\ Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} specifies that individual galaxy \pz s must have root-mean-square error $\sigma_z < 0.02 (1+z)$, $3 \sigma$ catastrophic outlier rate below $10\%$, and bias below $0.003$.
Specific science cases may have their own requirements on \pz\ performance that exceed those of the survey as a whole.
In that vein, the \lsst\ Dark Energy Science Collaboration (\lsstdesc) developed a separate SRD \citep{Mandelbaum:2018} that conservatively forecasts the constraining power of five cosmological probes, leading to even more stringent requirements on \pz\ performance, including those defined in terms of tomographically binned subsamples populations rather than individual galaxies.

@@

Though the standard has long been for each galaxy in a photometric catalogue to have a \pz\ point estimate and Gaussian error bar, even early applications of \pz s in precision cosmology indicate the inadequacy of point estimates \citep{mandelbaum:2008} to encapsulate the degeneracies resulting from the nontrivial mapping between broad band fluxes and redshift.
Far from a hypothetical situation, such degeneracies are real consequences of the same deep imaging that enables larger galaxy catalogue sizes.
The lower luminosity and higher redshift populations captured by deeper imaging introduce major physical systematics to \pz s, among them the Lyman break/Balmer break degeneracy, that did not affect shallower large area surveys like the Sloan Digital Sky Survey \citep[\textsc{SDSS},][]{York:00} and Two Micron All Sky Survey \citep[\textsc{2MASS},][]{Skrutskie:06}.

@@

There are numerous techniques for deriving \pzpdf s, yet no one method has been established as clearly superior.
Consistent experimental conditions enable the quantification if not isolation of their differences, which can be interpreted as a sort of \textit{implicit prior} imparted by the method itself.
Comprehensive comparisons of \pz\ methods have been made before; the \Pz\ Accuracy And Testing \citep[\textsc{PHAT},][]{Hildebrandt:10} effort focused on \pz\ point estimates derived from many photometric bands.
\citet{Rau:2015} introduced a new method for improving \pzpdf s using an ordinal classification algorithm.
\textsc{DES} compared several codes for \pz\ point estimates and a subset with \pzpdf\ information \citep{Sanchez:14} and examined summary statistics of \pzpdf s for tomographically binned galaxy subsamples \citep{Bonnett:16}.

This paper is distinguished from other comparisons of \pz\ methods by its focus on the evaluation criteria for \pzpdf s and interpretation thereof.
In the absence of simulated data drawn from known redshift distributions, the very concept of a ``true PDF'' for an individual galaxy is unavailable, and we must instead rely on measures of ensemble behaviour to characterize PDF quality (see \S\,\ref{sec:metrics} for further discussion).
We aim to perform a comprehensive sensitivity analysis of \pzpdf\ techniques in order to ultimately select those that will become part of the \lsst\ pipelines, as part of a key project of the Photometric Redshifts working group of the \lsstdesc, described in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}}.
In this initial study, we focus on evaluating the performance of \pzpdf\ codes using PDF-specific performance metrics in a formally controlled experiment with complete and representative prior information (template libraries and training sets) to set a baseline for subsequent investigations.
This approach probes how each code considered exploits the information content of the data versus prior information from template libraries and training sets.
